---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
Load Libraries
```{r LoadLibraries, echo=TRUE}
library(ggplot2)
library(plyr)

```

## Loading and preprocessing the data
Unzip and load file and convert date
```{r LoadData, echo=TRUE}
activity <- read.csv(unz("activity.zip","activity.csv"))
activity$date <- as.Date(activity$date)
```


## What is mean total number of steps taken per day?
Apply sum and ignore NA values, plot to Histogram, and find Mean and Median
```{r CalculateMean, echo=TRUE}
stepsperday <- tapply(activity$steps,activity$date,sum,na.rm=TRUE)
qplot(stepsperday, geom="histogram", binwidth=1000, main="Histogram of Steps Per Day")
mean(stepsperday)
median(stepsperday)
```

## What is the average daily activity pattern?
```{r CalculateDaily, echo=TRUE}
stepsperinterval <-ddply(activity,.(interval), summarize,MeanSteps= mean(steps, na.rm=TRUE))
qplot(interval, MeanSteps,data =stepsperinterval, geom="line",  main="Average Steps Per Interval")

```

Which Interval has the maximum number of mean steps

```{r CalculateMaxAverageSteps, echo=TRUE}
stepsperinterval[stepsperinterval$MeanSteps==max(stepsperinterval$MeanSteps),]
```

## Imputing missing values

How Many Missing (NA) Steps are in the data

```{r CalculateMissingSteps, echo=TRUE}
sum(is.na(activity$steps))
```

Create a dataset where the meanSteps for the interval is applied, and use to overwrite NA where applicable
```{r imputemissingsteps, echo=TRUE}

activityImputed <- merge(activity,stepsperinterval)
activityImputed[is.na(activityImputed$steps),"steps"] <-activityImputed[is.na(activityImputed$steps),"MeanSteps"]

```



Apply sum and ignore NA values, plot to Histogram, and find Mean and Median for imputed dataset
```{r CalculateMeanImputed, echo=TRUE}
stepsperdayimputed <- tapply(activityImputed$steps,activityImputed$date,sum,na.rm=TRUE)
qplot(stepsperdayimputed, geom="histogram", binwidth=1000, main="Histogram of Steps Per Day")
mean(stepsperdayimputed)
median(stepsperdayimputed)
```

How has the imputation affected the Mean and Median Values
```{r imputeddifference, echo=TRUE}

mean(stepsperdayimputed)-mean(stepsperday)
median(stepsperdayimputed)-median(stepsperday)
```
## Are there differences in activity patterns between weekdays and weekends?

Create a weekday/weekend factor and summarise 

```{r dayofweek, echo=TRUE}

activityImputed$wdaywend <- as.factor(ifelse(weekdays(activityImputed$date) %in% c("Saturday","Sunday"), "Weekend", "Weekday"))
wdaywendcomp <-ddply(activityImputed,.(interval,wdaywend), 
                    summarize,MeanSteps= mean(steps, na.rm=TRUE))
```

Create a panel line plot comparing Weekday and Weekend
```{r comparisonwdaywend, echo=TRUE}
qplot(interval, MeanSteps,data =wdaywendcomp, geom="line", color = wdaywend, 
    main="", xlab ="Interval", ylab = "Number of Steps")+
    theme(legend.position = "none") + facet_grid( .~ wdaywend)+ facet_wrap(~wdaywend, ncol=1)
```